A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Additional Methods

### 2020-01-28 writeUp/supplementary_materials_methods.tex
corrected detinova parity calculation in SOM

\section{Detinova dissection experiments}\label{sec:detinova}
\cite{detinova1962age} provided an alternative approach to estimate the lifespan of individual specimens, based on sampling the parous rate in a population (the proportion of adult females that have laid eggs) using dissection to determine parity of each specimen. In what follows, we follow the derivation given by \cite{davidson1954estimation}, to determine an expression for the proportion of nulliparous mosquitoes in terms of the gonotrophic cycle length $G$, here assumed the same for each cycle subsequent to an adult emerging. If a constant daily survival rate $p$ is assumed, the proportion of emerging adults surviving to become parous is $p^G$. If the rate of recruitment into the adult classes is constant over time, the proportion of nulliparous mosquitoes is,
%
\begin{equation}\label{eq:detinova}
\begin{aligned}
\mathbb{n} &= \frac{p^G}{p^G + p^{2G} + p^{3G} + \cdots}\\
&= p^G / \frac{p^G}{1-p^G}\\
&= 1-p^G.
\end{aligned}
\end{equation}
%
The daily survival rate can hence be calculated by inverting eq. (\ref{eq:detinova}): $p = \sqrt[G]{\mathcal{P}}$, where $\mathcal{P}=1-\mathbb{n}$ is the proportion parous sampled from the population. The distribution of mosquitoes dying each day follows a geometric distribution, with probability of dying on a given day equal to $1-p$. The mean lifespan is, hence, given by $1/(1-p)$.

>>>>

\section{Detinova dissection experiments}\label{sec:detinova}
\cite{detinova1962age} provided an alternative approach to estimate the lifespan of individual specimens, based on sampling the parous rate in a population (the proportion of adult females that have laid eggs) using dissection to determine parity of each specimen. In what follows, we follow the derivation given by \cite{davidson1954estimation}, to determine an expression for the mean lifespan of a mosquito in terms of gonotrophic cycles (all assumed to be of the same duration, $G$). If the probability of dying in a given cycle, $p$, is assumed constant (i.e. there is no age-dependent mortality), then the probability a mosquito completes $Y$ cycles, is geometrically distributed,
%
\begin{equation}
Pr(Y=k) = (1-p)^k p,
\end{equation}
%
where $k= 0,1,2,...$. This implies that the proportion of adults surviving to become parous is $Pr(Y\geq 1) = 1 - Pr(Y=0) = 1-p=\mathcal{P}$. Therefore, the proportion of parous mosquitoes in any given sample $i$ of wild-caught mosquitoes, $\mathcal{P}^{\{i\}}$ is a direct estimator of the probability of surviving a given cycle, $1-p$. The mean lifespan of a mosquito in terms of completed gonotrophic cycles is then given by $(1-p)/p= \mathcal{P}/(1-\mathcal{P})$.

If, instead, we choose chronological time as our measure of lifespan, the above analysis still holds, although with $q$ now representing the probability of dying on a particular day, meaning $(1-q)^G$ is the probability of surviving a cycle. The daily survival probability $1-q$ is, hence, equal to $\sqrt[G]{\mathcal{P}}$, and the mean lifespan is given by $q/(1-q)=\sqrt[G]{\mathcal{P}}/(1-\sqrt[G]{\mathcal{P}})$.
